---
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Name:__


# For 03/17/16

## 4E1
line one

## 4E2
two (Mu and Sigma)

## 4E3
Posterior=(Likelihood X Prior)/Average Likelihood

pr(μ,σ|y) = ∏i Normal(yi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10) / ∫∫ ∏i Normal(yi|μ, σ)Normal(μ|0, 10)Uniform(σ|0, 50)dμdσ

## 4M1

```{r}
library("rethinking")
sample_mu <- rnorm( 1e4 , 0 , 10 )
sample_sigma <- runif( 1e4 , 0 , 10 )
prior_y <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_y )
```

## 4M2

```{r}
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]


flist <- alist(
height ~ dnorm( mu , sigma ) ,
mu ~ dnorm( 0 , 10 ) ,
sigma ~ dunif( 0 , 10 )
)
m4.1 <- map( flist , data=d2 )
precis( m4.1 )
```

# For 03/24/16

## 4E4
line 2
## 4E5
Three: α,β,and σ

## 4M3
yi ~Normal(μ,σ)

μi= α + βxi

α~Normal(0, 50)

β~Normal(0, 10)

σ~Uniform(0, 50)

## 4M4
hi=height
xi=year

hi ~Normal(μi,σ)

μi= α + βxi

α~Normal(178, 100)

β~Normal(0, 10)

σ~Uniform(0, 50)

## 4M5
hi ~Normal(μi,σ)

μi= α + βxi

α~Normal(120, 100)

β~Normal(10, 10)

σ~Uniform(0, 50)

Center α around the mean given from the first year instead.

Assume that there is some positive effect between year and height, so change from a 0 prior for β's mean to 10.

## 4M6
hi ~Normal(μi,σ)

μi= α + βxi

α~Normal(120, 64)

β~Normal(10, 10)

σ~Uniform(0, 50)

I would change the prior for α's variance to 64.

# For 03/31/16

## 4H1

```{r}
# load data
library(rethinking)
data(Howell1)
d <- Howell1

# fit model linear regression
m1 <- map(
alist(
height ~ dnorm( mu , sigma ) ,
mu <- a + b*weight ,
a ~ dnorm( 156 , 100 ) ,
b ~ dnorm( 0 , 10 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data=d)

#plot
plot( height ~ weight , data=d )
abline( a=coef(m1)["a"] , b=coef(m1)["b"] )

#sample

post <- extract.samples(m1)

#Prediction
weight.seq <- c(46.95, 43.72, 64.78, 32.59, 54.63)
mu <- link( m1 , data=data.frame(weight=weight.seq) )

mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
mu.mean
mu.HPDI[,1]
mu.HPDI[,2]
mu.HPDI[,3]
mu.HPDI[,4]
mu.HPDI[,5]
```


## 4H2

#A

```{r}
# load data
library(rethinking)
data(Howell1)
d <- Howell1
d2=d[d$age < 18, ]
nrow(d2)

# fit model linear regression
m2 <- map(
alist(
height ~ dnorm( mu , sigma ) ,
mu <- a + b*weight ,
a ~ dnorm( 100 , 100 ) ,
b ~ dnorm( 0 , 10 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data=d2)

#estimates
precis( m2 )
```

The slope is parameter b in the table. For every 10 units of increase in weight a child gets about 27cm taller.

#B

```{r}
#sample
weight.seq <- seq(0, 40, by=1)

post <- extract.samples(m2)

mu <- link( m2 , data=data.frame(weight=weight.seq) )

mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )


#plot
plot( height ~ weight , data=d2 , col=col.alpha(rangi2,0.5) )
lines( weight.seq , mu.mean )
shade( mu.HPDI , weight.seq )

sim.height <- sim(m2, data = list(weight=weight.seq))
height.PI <- apply(sim.height, 2, PI, prob=0.89)
shade(height.PI, weight.seq)
```

#C

The model might be better fit with a curve instead of linear regression.

## 4H3

#A

```{r}
# load data
library(rethinking)
data(Howell1)
d <- Howell1

# fit model linear regression
logweight=log(d$weight)

m3 <- map(
alist(
height ~ dnorm( mu , sigma ) ,
mu <- a + b*logweight ,
a ~ dnorm( 156 , 100 ) ,
b ~ dnorm( 0 , 10 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data=d)

precis(m3)
#plot
plot( height ~ logweight , data=d )


```

#B

```{r}
plot( height ~ weight , data=Howell1 ,
col=col.alpha(rangi2,0.4) )


```

