---
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Name:__


# For 03/17/16

## 4E1
line one

## 4E2
two (Mu and Sigma)

## 4E3
Posterior=(Likelihood X Prior)/Average Likelihood

pr(μ,σ|y) = ∏i Normal(yi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10) / ∫∫ ∏i Normal(yi|μ, σ)Normal(μ|0, 10)Uniform(σ|0, 50)dμdσ

## 4M1

```{r}
library("rethinking")
sample_mu <- rnorm( 1e4 , 0 , 10 )
sample_sigma <- runif( 1e4 , 0 , 10 )
prior_y <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_y )
```

## 4M2

```{r}
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]


flist <- alist(
height ~ dnorm( mu , sigma ) ,
mu ~ dnorm( 0 , 10 ) ,
sigma ~ dunif( 0 , 10 )
)
m4.1 <- map( flist , data=d2 )
precis( m4.1 )
```

# For 03/24/16

## 4E4
line 2
## 4E5
Three: α,β,and σ

## 4M3
yi ~Normal(μ,σ)

μi= α + βxi

α~Normal(0, 50)

β~Normal(0, 10)

σ~Uniform(0, 50)

## 4M4
hi=height
xi=year

hi ~Normal(μi,σ)

μi= α + βxi

α~Normal(178, 100)

β~Normal(0, 10)

σ~Uniform(0, 50)

## 4M5
hi ~Normal(μi,σ)

μi= α + βxi

α~Normal(120, 100)

β~Normal(10, 10)

σ~Uniform(0, 50)

Center α around the mean given from the first year instead.

Assume that there is some positive effect between year and height, so change from a 0 prior for β's mean to 10.

## 4M6
hi ~Normal(μi,σ)

μi= α + βxi

α~Normal(120, 64)

β~Normal(10, 10)

σ~Uniform(0, 50)

I would change the prior for α's variance to 64.

# For 03/31/16

## 4H1

## 4H2

## 4H3
